Mathematical Language Processing \\ Project

### 2014-02-27 paper.tex
wrapped textblocks to be more readable in text editors

\begin{abstract}
In this paper, we present a solution for automated definition discovery for the variables occurring in mathematical formulae.
Ordinary natural language processing approaches that are trained based on large, manually annotated news corpora fall short in reasonable quality of definition discovery for scientific texts.
In the \emph{Mathematical Language Processing} project, we discover definitions %we did not discover any relations
% a relation would be something like "let <subj> be a <object>", or "we call <subj> <obj>", or "<subject> is <object>" or just "<subj> <obj>"
 for identifiers in formulae from the surrounding text by statistical methods.
 %with a machine learning approach for classification. This was also not done in this paper
%We extract the identifiers from the formulae and track co-occurrences of those in the surrounding text.
The evaluation of our prototypical system, applied on the Wikipedia text corpus shows that our approach augments the user experience for articles with mathematical formula substantially.
While hovering the identifiers occurring in the equation a pop-up with the most probable explanations for the identifiers occurs. The displayed definition information provides a good match to the actual meaning of the identifiers.
\keywords{machine learning, text mining, parallel computing}
\end{abstract}

@@

\subsection{Related Work}
\citeauthor{Quoc2010} (2010) proposed in \cite{Quoc2010} an approach for relating whole formulas to sentences and their describing paragraphs. \citeauthor{Yokoi} (2010) trained a \emph{support vector machine} to extract natural language descriptions for mathematical expressions \cite{Yokoi}.

@@

The two entities have some characteristics that can be used for filtering. First of all, the relevant variables can be retrieved accurately from \TeX\ formatted formulas (see Section \ref{vr}). The fact that a relation consists of two entities and we know that one must be the variable, leads to an obvious conclusion: Sentences without a variable cannot contain a definition relation and can be discarded.
Second, one can assume that the other entity must be a noun or a compound noun. Thus, all words that are not nouns or adjectives describing these nouns, can also be discarded.

@@

Furthermore, we take distributional properties into account. The classic tf-idf \cite{Salton86} statistic reflects the importance of a term to a document. Unfortunately, the \emph{inverse document frequency} is not applicable to definition relations. The frequency distribution of all terms, that participate in various relations is unknown. Thus, using tf-idf could affect the sensitivity of our rating statistic significantly. E.g., `Hamiltonian' as well as `length' are both valid entities. Obviously, the number of documents containing one of these words, within a corpus like Wikipedia, differs a lot.

@@

\tikzset{actor/.style={
        rectangle,
        minimum size=6mm,
        very thick,
        draw=gray!50!black!50,
        top color=white,
        bottom color=gray!50!black!20
    },
    arrow/.style={
        -latex, thick, shorten <=2pt,shorten >=2pt
    }
}
\begin{figure}[H]
	\begin{tikzpicture}[node distance=5mm and 8mm]
		\node (Input) [align=center]{Wiki Dumps};
		\node (DocumentParser) [actor, right=of Input, align=center] {\emph{Map}\\\textbf{Parser}};
		\node (Candidates) [actor, right=of DocumentParser, align=center] {\emph{CoGroup}\\\textbf{Kernel}};
		\node (Sentence) [actor, above=of DocumentParser, align=center] {\emph{Map}\\\textbf{Tagger}};
		\node (Filter) [actor, right=of Candidates, align=center] {\emph{Reduce}\\\textbf{Filter}};
		\node (Output) [right=of Filter,align=center]{Raw\ Candidates};
		\draw[arrow] (Input)--(DocumentParser);
		\draw[arrow] (DocumentParser)--(Sentence);
		\draw[arrow] (DocumentParser)--(Candidates);
		\draw[arrow] (Sentence.east)--(Candidates.north);
		\draw[arrow] (Candidates)--(Filter);
		\draw[arrow] (Filter)--(Output);
	\end{tikzpicture}
\caption{Data flow of the PACT program}
\end{figure}
There are three computation heavy parts in the processor. First, the raw Wikipedia document dumps need to be transformed into plain text. At the same step, the formulas need to be evaluated and replaced by place-holder words. We used the \emph{WikiText} framework of the Mylyn project with a custom builder to transform the document into plain text. For analysing the \TeX \ mark-up in the \texttt{<math/>} tags, \emph{SnuggleTeX} transforms the formulas into their MathML representations, where variable names can be easily extracted.

>>>>

\begin{abstract}
In this paper, we present a solution for automated definition discovery for
the variables occurring in mathematical formulae. Ordinary natural language
processing approaches that are trained based on large, manually annotated news
corpora fall short in reasonable quality of definition discovery for
scientific texts. In the \emph{Mathematical Language Processing} project, we
discover definitions
%we did not discover any relations  a relation would be something like "let
%<subj> be a <object>", or "we call <subj> <obj>", or "<subject> is <object>"
%or just "<subj> <obj>"
for identifiers in formulae from the surrounding text by statistical methods.
%with a machine learning approach for classification. This was also not done
%in this paper We extract the identifiers from the formulae and track co-
%occurrences of those in the surrounding text.
The evaluation of our prototypical system, applied on the Wikipedia text
corpus shows that our approach augments the user experience for articles with
mathematical formula substantially. While hovering the identifiers occurring
in the equation a pop-up with the most probable explanations for the
identifiers occurs. The displayed definition information provides a good match
to the actual meaning of the identifiers.
\keywords{machine learning, text mining, parallel computing}
\end{abstract}

@@

\subsection{Related Work}
\citeauthor{Quoc2010} (2010) proposed in cite{Quoc2010} an approach for
relating whole formulas to sentences and their describing paragraphs.
\citeauthor{Yokoi} (2010) trained a emph{support vector machine} to extract
natural language descriptions for mathematical expressions cite{Yokoi}.

@@

The two entities have some characteristics that can be used for filtering.
First of all, the relevant variables can be reieved accurately from \TeX\
formatted formulas (see Section \ref{vr}). The fact that a relation consists
of two entities and we know that one must be the variable, leads to an obvious
conclusion: Sentences without a variable cannot contain a definition relation
and can be discarded. Second, one can assume that the other entity must be a
noun or a compound noun. Thus, all words that are not nouns or adjectives
describing these nouns, can also be discarded.

@@

Furthermore, we take distributional properties into account. The classic tf-
idf \cite{Salton86} statistic reflects the importance of a term to a document.
Unfortunately, the \emph{inverse document frequency} is not applicable to
definition relations. The frequency distribution of all terms, that
participate in various relations is unknown. Thus, using tf-idf could affect
the sensitivity of our rating statistic significantly. E.g., `Hamiltonian' as
well as `length' are both valid entities. Obviously, the number of documents
containing one of these words, within a corpus like Wikipedia, differs a lot.

@@

\tikzset{actor/.style={
        rectangle,
        minimum size=6mm,
        very thick,
        draw=gray!50!black!50,
        top color=white,
        bottom color=gray!50!black!20
    },
    arrow/.style={
        -latex, thick, shorten <=2pt,shorten >=2pt
    }
}
\begin{figure}[H]
	\begin{tikzpicture}[node distance=5mm and 8mm]
		\node (Input) [align=center]{Wiki Dumps};
		\node (DocumentParser) [actor, right=of Input, align=center] {\emph{Map}\\\textbf{Parser}};
		\node (Candidates) [actor, right=of DocumentParser, align=center] {\emph{CoGroup}\\\textbf{Kernel}};
		\node (Sentence) [actor, above=of DocumentParser, align=center] {\emph{Map}\\\textbf{Tagger}};
		\node (Filter) [actor, right=of Candidates, align=center] {\emph{Reduce}\\\textbf{Filter}};
		\node (Output) [right=of Filter,align=center]{Raw\ Candidates};
		\draw[arrow] (Input)--(DocumentParser);
		\draw[arrow] (DocumentParser)--(Sentence);
		\draw[arrow] (DocumentParser)--(Candidates);
		\draw[arrow] (Sentence.east)--(Candidates.north);
		\draw[arrow] (Candidates)--(Filter);
		\draw[arrow] (Filter)--(Output);
	\end{tikzpicture}
\caption{Data flow of the PACT program}
\end{figure}

There are three computation heavy parts in the processor. First, the raw
Wikipedia document dumps need to be transformed into plain text. At the same
step, the formulas need to be evaluated and replaced by place-holder words. We
used the \emph{WikiText} framework of the Mylyn project with a custom builder
to transform the document into plain text. For analysing the \TeX \ mark-up in
the \texttt{<math/>} tags, \emph{SnuggleTeX} transforms the formulas into
their MathML representations, where variable names can be easily extracted.
